---
title: "ENVS-193DS-FINAL"
author: "Damian Rojas"
format: pdf
---


```{r}
#package set up
library(tidyverse)
library(janitor)
library(readxl)
library(DHARMa)
library(ggeffects)
nests <- read.csv("nest_data_final.csv")
walks <- read.csv("walks_data2.csv")
```

# 1
## a)
In part 1 they used a binomial (logistical regression) test. In part 2, they used a chi-square test to check for a relationship. 

## b)
*Part 1* : With a logistical regression we find that distance to water edge does  predict the probability of American Avocet microhabitat use. Water edge proximity is an important variable American Avocet's use to determine where they live. 

*Part 2*: We found that dependening on bird species, preference in habitat type also changes using a Chi-Square test. Habitat type preference is not consistent among varying bird species. 

## c)
One additional variable that could influence the probability of American Avocet microhabitat use is temperature. Temperature is a continuous variable (measured in either degrees celsius or fahrenheit) and could be indicative of microhabitat use along with water edge proximity as typically areas close to water have much more temperate climates, which could be a preference of the American Avocet. 

# 2
## a)

```{r}
#| echo: false
#cleaning and wrangling
nests_clean <- nests |> 
  clean_names() |> 
  select(height_cm, ant_species, case_control) |> 
  filter(ant_species %in% c("AB", "RRB", "BBR")) |> 
  mutate( ant_species = factor(ant_species, 
                               levels = c("AB", "RRB", "BBR"),
                               labels = c(
                                 "C. sjostedi", "C. mimosae", "C. nigriceps"
                               ))) |> 
mutate(scientific_name = case_when(
  ant_species == "C. sjostedi" ~ "Crematogaster sjostedi", ant_species == "C. mimosae" ~ "Crematogaster mimosae", 
  ant_species == "C. nigriceps" ~ "Crematogaster nigriceps"
  
))

nests_clean |> 
slice_sample(n = 10)
```

## b)

The 1s and 0s show whethere or not a tree contains a nest (1) or is available for nesting (0)

## c)
Tree height is a continuous variable. As tree height increases the probability of nest occurrence would increase because it provides more areas for birds to create nests in the area. (unabale to access paper cannot cite this information, merely a guess)

Acacia Ant species is a categorical variable. The species of Acacia ant can fend off insects that harm bird offspring which increases the probability of nest occurance. (abstract, paragraph 1). 

## d)
```{r}
model_tabel <- tibble(
  model_number = c(1,2), 
  predictors = c("height_cm + ant_species",
                 "height_cm * ant_species") , 
  interaction_term = c("No", "Yes"),
  model_description = c("Contains tree height and ant species main effects", "Contains interaction between btree height and ant species")
)
```

## e)
```{r}
model0 <- lm(data = nests_clean,
  case_control ~ height_cm + ant_species
)

model1 <- lm(data = nests_clean, 
             case_control ~ height_cm*ant_species)
```

## f)

```{r}
AIC(model0,model1)
```

The best model as determined by Akaike's Information Criterion was the model predicting whether a tree was nested or not from tree height and ant species with an interaction term. 

## g)
```{r}
par(mfrow = c(2,2))
plot(model1)

```

## h)

```{r}

pred <- ggpredict(model1, terms = c("height_cm", "ant_species")) |> 
  mutate

ggplot(pred, aes (x = x, y = predicted, color = group,)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.3) +
  #googled how to keep predictions within 0-1 range due to binary restrictions
  coord_cartesian(ylim = c(0, 1)) +
  geom_jitter(data = nests_clean, 
              aes(x = height_cm, y = case_control), 
              color = "red4", 
              height = 0) +
  facet_wrap (~ group) +
  theme_bw() +
  theme(legend.position = "none")
```

## i)
**Figure 1** : Relationship between ant species and nested trees; 
nested trees represented as nested(1) and available (0) by red points, with ant species split and represented by green confidence interval, green confidnece interval, and blue confidence interval. 
Data citation: Lujan, E., Nielsen, R., Short, Z., Wicks, S., Nderitu Watetu, W., Malingati Khasoha, L., Palmer, T. M., Goheen, J. R., & Alston, J. M. (2023). Data, code, and metadata for: Symbiotic acacia ants drive nesting behavior by birds in an African savanna [Data set]. Zenodo. https://doi.org/10.5281/zenodo.8373322

## j)

```{r}
#create new object, use ggpredict and model and set height to max height 600cm for prediction. 
max_height_pred <- ggpredict(model1, terms = c("height_cm [600]", "ant_species"))
max_height_pred
```

## k) 

The predicted probability of nest site occupancy at 600cm changes across ant species with the greatest probability lying with C. nigriceps, the second greatest with C. mimosae, and the lowerst probability with C. sjostedi. With the figures in part h, we can see that C. sjostedi ants typically are found in shorter trees, and C. nigriceps typically being located in taller trees. Certain acacia ant species (C. mimosae and C. nigriceps) are more likely to be found at trees with bird nests, due to their ability to fend off insects that can harm bird offspring. 

# 3
## a)
The visualizations are different because the affective visualization tells a story about the data
whereas the visualizations in homework 2 show just the raw data, no story. 
Within the visualizations the only pattern I can see between the both of them is that they don't really vary, the data remains stable with few outliers in between. 
During week 9 workshop I hadn't created a visualization because I had submitted an extension for homework 3, so during the classmate feedback i bounced ideas off my classmates and they said that showing the route I would take was a very solid idea so I launched my visualization idea off of that interaction. 

## b)
Based on my test of walk time and assignment workload, I believe a linear regression would be an appropriate test because 
workload was measured numerically and walk time is a continuous variable, which can be predicted with the linear regression. 

## c)

```{r}
modelw <- lm(data = walks, walk_time_s ~ work_due)

plot(modelw)

modelw
```

## d)

```{r}
walk_predictions <- ggpredict(modelw, terms = c("work_due")) 

ggplot(walk_predictions, aes (x = x, y = predicted)) +
  geom_point(data = walks, aes(x = work_due, y = walk_time_s), color = "green4") +
  geom_ribbon(aes(ymin =conf.low, ymax = conf.high), alpha = 0.3) +
  geom_line(color = "blue2", linewidth = 1) +
  theme_bw() +
  labs(x = "Assignments due at time of walk (#)",
       y = "Length of walk (seconds)")


```

## e)
**Figure 2** : Impact of workload on walk time; Blue line represents relationship between length of walk in seconds and work due, showing how with more work due the shorter the walk is. Represented in green points are the observed data with a 95% CI ribbon surrounding the data. 

## f) 
The linear regression model for the relationship of work due and walk time over the course of the quarter (n = 16). The model shows that when 0 work is due, the predicted length of the walk is 882.47 seconds. For each assignment due this overall time is reduced by 9.25 seconds. This shows that there is a negative relationship between work due and length of walk. 
